---
title: 'Reproducibe Research: Peer Assessment 1'
author: "Mohammed Teslim"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
```

### Loading and Processing Data

```{r Loading}
if(!dir.exists("RepData_PeerAssessment1")){
    dir.create("RepData_PeerAssessment1")
}

if(!str_ends(getwd(), "/RepData_PeerAssessment1")){
    setwd("/RepData_PeerAssessment1")
}

activity_data <- read.csv("activity.csv")
    activity_data$date <- as.Date(activity_data$date)
    
```

### Mean total number of steps taken per day

```{r mean, warning=FALSE, comment=""}
steps_per_day <- activity_data %>% group_by(date) %>% summarise(TotalSteps = sum(steps,na.rm = T))

ggplot(steps_per_day) + geom_histogram(aes(TotalSteps), fill = "red", color = "blue", binwidth = 1000)

mean_steps_per_day <- mean(steps_per_day$TotalSteps)

median_steps_per_day <- median(steps_per_day$TotalSteps)

print(paste("Mean of the total number of steps equals", as.integer(mean_steps_per_day)))

print(paste("Median of the total number of steps per day", as.integer(median_steps_per_day)))

```

### Average Daily Activity Pattern

```{r Average, comment=""}
Average_Steps <- activity_data %>% group_by(date) %>% summarise(Daily_Average = mean(steps, na.rm = T))

ggplot(Average_Steps, aes(date, Daily_Average)) + geom_line(color = "#69b3a2")

print(paste("Maximum number of steps in a 5-minute interval is", as.integer(max(Average_Steps$Daily_Average, na.rm = T))))
```

### Imputing missing values

There are 8 date observations that have NA as missing values. The steps variable for these dates are missing all through, these implies that there are no steps recorded for those dates. Therefore, there is nothing to be used for those dates (i.e either mean or median) to fill in the missing values. As a result of this, the mean of the average daily activity computed above will be used as the input for these missing values. 
```{r missing values, warning=FALSE, comment=""}
num_missing <- sum(is.na(activity_data$steps))
print(paste("Total number of missing data is", num_missing))

# Calculate the mean of the daily averages
overall_mean <- as.integer(mean(Average_Steps$Daily_Average, na.rm = T))
activity_data2 <- activity_data
activity_data2$steps[is.na(activity_data2$steps)] <- 37

steps_per_day2 <- activity_data2 %>% group_by(date) %>% summarise(TotalSteps2 = sum(steps,na.rm = T))

ggplot(steps_per_day2) + geom_histogram(aes(TotalSteps2), fill = "red", color = "blue", binwidth = 1000)

mean_steps_per_day2 <- mean(steps_per_day2$TotalSteps2)

median_steps_per_day2 <- median(steps_per_day2$TotalSteps2)

print(paste("Mean of the total number of steps equals", mean_steps_per_day2))

print(paste("Median of the total number of steps per day", median_steps_per_day2))
```

These values can be seen to differ from the estimates of the first assignment but only in a little way.

### Differences in activity patterns between weekdays and weekends

```{r weekday/ends }
activity_data2 %>% mutate(Day_type = wday(date, label = T)) %>% mutate(Day_type = fct_collapse(Day_type, Weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"), Weekend = c("Sat", "Sun"))) %>% group_by(Day_type,date) %>% summarize(avesteps = mean(steps)) %>% ggplot(aes(date, avesteps)) + geom_line() + facet_wrap(~ Day_type)
```